---
Title: Exploratory Graphs
Instructor: Roger D. Peng, *JHU via Coursera*
author: "Krystella Rattan"
date: "5/30/2022"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Exploratory Graphs are used:

* To understand data properties
* To find patterns in data
* To suggest modeling strategies
* To "debug" analyses

Characteristics of exploratory graphs:

* They are made quickly
* A large number are made
* The goal is for personal understanding
* Axes/legends are generally cleaned up later
* Colour/size are primarily used for information

>> Exploratory graphs are meant to familiarise the analyst with the data

## **Example: Ambient Air Pollution in the US**

**Data:** Annual average PM 2.5 averaged over the period 2008 through 2010

**Data Source:** https://github.com/jtleek/modules/blob/master/04_ExploratoryAnalysis/exploratoryGraphs/data/avgpm25.csv

**Context:**
The US Environmental Protection Agency (EPA) sets the national ambient air quality standards for outdoor air pollution

For fine particle pollution (PM2.5), the "annual mean, averaged over 3 years" cannot exceed 12 $\mu$g / $m^{3}$

**Question:**
Are there any counties in the US that exceed that national standard for fine particle pollution?

```{r}
knitr::opts_chunk$set(echo = TRUE)

pollution <- read.csv("E:/Work/Data Science/R/4 Exploratory Data Analysis/avgpm dataset/avgpm25.csv", colClasses = c("numeric", "character", "factor", "numeric", "numeric"))

```

```{r}
knitr::opts_chunk$set(echo = TRUE)

head(pollution)

```


*Do any counties exceed the standard of 12 $\mu$g / $m^{3}$ ?*


### One-Dimensional Summaries of Data

**Five-number summary**
```{r}
summary(pollution$pm25)

```

---

**Boxplots**
```{r}
boxplot(pollution$pm25, col = "blue")

```

---

**Histograms**
```{r}
hist(pollution$pm25, col = "green")

```
```{r}
hist(pollution$pm25, col = "green")
rug(pollution$pm25)

```


```{r}
hist(pollution$pm25, col = "green", breaks = 100)
rug(pollution$pm25)

```


---

**Overlaying features**
```{r}
boxplot(pollution$pm25, col = "blue")
abline(h = 12)

```

```{r}
hist(pollution$pm25, col = "green")
abline(v = 12, lwd = 2)
abline(v = median(pollution$pm25), col = "magenta", lwd =4)

```


---

**Barplot**

```{r}
barplot(table(pollution$region), col = "wheat", main = "Number of Counties in Each Region")

```





























